Palomar 5 Gaps

### 2024-09-17 paper/main.tex
updated equations for collision and 1D profile

\documentclass{aa}

@@

      To construct the 1D density profiles, we first bin the data using the $\sqrt{N}$-rule for determining the number of bins, where N is the number of data points which is 10$^5$. After histogramming the 1D profiles, we apply a boxcar smoothing technique, where each central value is replaced by the median of the surrounding XX adjacent data points, corresponding to a smoothing length of approximately 300 parsecs. This procedure reduces high-frequency noise and smooths the profiles, as shown in Fig.~\ref{fig:profiles}, where the center of mass has been effectively removed.

      With the smoothed 1D density profiles in hand, we consider the \textit{vanilla} simulations to be truth, and search for locations that are significantly less dense than \textit{vanilla} in a way that is greater than stochastic chance. To do so, we first compute the log ratio of the two profiles:
      \begin{equation}
        \mathcal{R} = \log\left(\frac{\rho_f}{\rho_v}\right),
        \end{equation}
      where $\rho$ is the 1D density, $v$ corresponding to the vanilla simulation, and $f$ corresponding to the full simulation, and $R$ is the log ratio of the two. Using the log ratio is important, because the distribution is symmetric and zero centered. For example if we only used the ratio, then the codomain would be $\frac{\rho_f}{\rho_v} \in [1,+\inf]$ for $\rho_f > \rho_v$, and $\frac{\rho_f}{\rho_v} \in [0,1]$ for $\rho_f < \rho_v$.

      Yet, we only wish to consider bins with a significant amount of counts. We find this threshold by asserting that the signal to noise ratio must be greater than 50. The signal is based on the \textit{vanilla} profile, and the noise is the error propogation given poisson uncertainties for the number of counts a given bin: 
      \begin{equation}
        \mathcal{S}/\mathcal{N} =50= \frac{\log{N}}{\frac{1}{\ln(10)} \sqrt{N}}
        \end{equation}

      we compute the difference between the logarithms of the binned counts for each profile and search for points that fall more than two standard deviations below the mean. However, we only consider regions that surpass a predefined significance threshold, where the significance is determined by the uncertainties in the binned counts governed by Poisson statistics multiplied by a factor of 50. This noise floor was chosen to eliminate the low-density regions near Palomar 5, where gaps are more likely caused by epicyclic oscillations rather than impacts from globular clusters. The identified gap candidates are marked and labeled on the plots.

      This method, however, is not without limitations, particularly when detecting smaller gaps. As discussed in Erkal et al. (2015), a gap's size is not necessarily indicative of a weak impact but could result from a more recent interaction. Additionally, because our streams have finite width, some gaps may be oriented obliquely relative to the stream axis. In these cases, marginalizing over y' can dilute the gap's signal, making it difficult to detect. An example of this is the small gap around 5 kpc, which, as we will demonstrate later, is caused by the interaction with NGC 2808 and appears in most simulations.

@@

    In this refinement process, we aim to pinpoint both the exact location of the impact along the stream and the precise moment it occurred. To do so, we fit a third-order parametric polynomial to the stream, using the saved snapshots from our simulations.
    \begin{align}
      x(\tau) &= a_0 + a_1 \tau + a_2 \tau^2 + a_3 \tau^3, \\ 
      y(\tau) &= b_0 + b_1 \tau + b_2 \tau^2 + b_3 \tau^3, \\
      z(\tau) &= c_0 + c_1 \tau + c_2 \tau^2 + c_3 \tau^3,
      \end{align}
    where $x$, $y$, and $z$ represent the parametric line describing the stream in galactocentric coordinates, $\tau$ is the stream coordinate in time as described in the previous section, and is used as the independent variable in this analysis. The coefficients $a_i$, $b_i$, and $c_i$ are the polynomial coefficients. We found that a second-order polynomial was insufficient to capture the curvature along the full length of the stream, with divergence from the star particle positions occurring at the extremes. A third-order polynomial was sufficient and desirable, as it is the lowest order that adequately captures the path of the stream over the entire length under consideration.

@@

    \begin{align}
      x(t,\tau) &= a_0(t) + a_1(t)\tau + a_2(t) \tau^2 + a_3(t)\tau^3, \\ 
      y(t,\tau) &= b_0(t) + b_1(t)\tau + b_2(t) \tau^2 + b_3(t)\tau^3, \\
      z(t,\tau) &= c_0(t) + c_1(t)\tau + c_2(t) \tau^2 + c_3(t)\tau^3.
      \end{align}
    The values of the coefficients as a function of time are obtained through linear interpolation, ensuring that the coefficients at the snapshot times match the values constrained by the simulation data.

@@

    \begin{align}
      \Delta \vec{v} &=\vec{v_p} - \vec{v_s} \\
      w_\parallel &= \left(\Delta \vec{v}\right)\cdot \hat{v_s}\\  
      w_\perp &=  \sqrt{\Delta v ^2 - w_\parallel ^ 2}
      \end{align}

>>>>

\documentclass[draft]{aa}

@@

      To construct the 1D density profiles, we first bin the data using the $\sqrt{N}$-rule for determining the number of bins, where N is the number of data points which is 10$^5$. After histogramming the 1D profiles, we apply a boxcar smoothing technique, where each central value is replaced by the median of the surrounding XX adjacent data points, corresponding to a smoothing length of approximately XX parsecs. This procedure reduces high-frequency noise and smooths the profiles, as shown in Fig.~\ref{fig:profiles}, where the center of mass has been effectively removed.

      With the smoothed 1D density profiles in hand, we consider the \textit{vanilla} simulations to be truth, and search for locations that are significantly less dense than \textit{vanilla} in a way that is greater than stochastic chance. To do so, we first only consider bins that are significant, to do this we impose a signal to noise ratio threshold, $\mathcal{SNR}$, of $\mathcal{SNR} = 5$. We assert that the singal is the log of the counts per bin of the vanilla 1D density profile, and we propogate the errors accordingly assert that the error in the counts is poisson distribution. As such, we are able to find a counts threshold, and discard any bins with out at least this number of counts. This $N$ is given by the transandental equation:
      \begin{equation}
        \mathcal{SNR} = \ln(10) \log_{10}\left(N_v\right)\sqrt{N_v}. 
        \end{equation}
      By setting a threshold of $\mathcal{S/R} = 5$, we solve for the equation by using \texttt{scipy.optimize.fsolve} and find that N must be greater than 7.

      After discarding the insignificant bins, we compute the log ratio of the counts between the vanilla and full simulations: 
      \begin{equation} 
        \mathcal{R}_i = \log_{10}\left(\frac{N_{f,i}}{N_{v,i}}\right),
        \end{equation}
      where $\mathcal{R}$ is the log ratio, $N_f$ are the number of counts from the full simulation, and $i$ is the index for a given bin. Subsequently, we inspect the resultant distribution of $\mathcal{R}_i$. This distribution would be gaussian if the differences between the density profiles are due to the sum of stochastic processes similar in magnitude, as in the central limit theorem. Therefore, we flag all locations that are underdense by two standard deviations.

      This method, however, is not without limitations, particularly when detecting smaller gaps. As discussed in Erkal et al. (2015), a gap's size is not necessarily indicative of a weak impact but could result from a more recent interaction. Additionally, because our streams have finite width, some gaps may be oriented obliquely relative to the stream axis. In these cases, marginalizing over y' can dilute the gap's signal, making it difficult to detect. As will be discussed in the results, this is the case for most of the gaps caused by NGC2808. Thus, studying the differences between the two 1D density profiles is an \textit{aid} to assist in visual inspect.

@@

    In this refinement process, we aim to pinpoint both the exact location of the impact along the stream and the precise moment it occurred. To do so, we fit a third-order parametric polynomial to the stream, using the saved snapshots from our simulations.

    \begin{equation}
      \vec{s}(\tau) = 
      \left\{
        \begin{aligned}
          x(\tau) &= a_0 + a_1 \tau + a_2 \tau^2 + a_3 \tau^3 \\ 
          y(\tau) &= b_0 + b_1 \tau + b_2 \tau^2 + b_3 \tau^3 \\
          z(\tau) &= c_0 + c_1 \tau + c_2 \tau^2 + c_3 \tau^3
        \end{aligned}
      \right.
  \end{equation}  
    where $x$, $y$, and $z$ represent the parametric line describing the stream in galactocentric coordinates, $\tau$ is the stream coordinate in time as described in the previous section, and is used as the independent variable in this analysis. The coefficients $a_i$, $b_i$, and $c_i$ are the polynomial coefficients. We found that a second-order polynomial was insufficient to capture the curvature along the full length of the stream, with divergence from the star particle positions occurring at the extremes. A third-order polynomial was sufficient and desirable, as it is the lowest order that adequately captures the path of the stream over the entire length under consideration.

@@

    \begin{equation}
      \vec{s}(t,\tau) = 
      \left\{
      \begin{aligned}
        x(t,\tau) &= a_0(t) + a_1(t)\tau + a_2(t) \tau^2 + a_3(t)\tau^3 \\ 
        y(t,\tau) &= b_0(t) + b_1(t)\tau + b_2(t) \tau^2 + b_3(t)\tau^3 \\
        z(t,\tau) &= c_0(t) + c_1(t)\tau + c_2(t) \tau^2 + c_3(t)\tau^3.
        \end{aligned}
      \right.
    \end{equation}
    The values of the coefficients as a function of time are obtained through linear interpolation, ensuring that the coefficients at the snapshot times match the values constrained by the simulation data.

@@

    \begin{equation}
        \begin{aligned}
          \Delta \vec{v} &=\vec{v_p} - \vec{v_s} \\
          w_\parallel &= \left(\Delta \vec{v}\right)\cdot \hat{v_s}\\  
          w_\perp &=  \sqrt{\Delta v ^2 - w_\parallel ^ 2}
          \end{aligned}
        \end{equation}
